\ab{Constraints on quasi--normal-mode frequencies with LIGO-Virgo binary--black-hole observations}

### 2021-03-18 papers/intro_paper/discussion.tex
changes in Sec. V

These constraints could in principle be used to constrain specific
non-GR theories and exotic compact objects (see
e.g.~\cite{Glampedakis:2017cgd,Cardoso:2019rvt,Maggio:2020jml}). However,
QNM computations in beyond-GR theories have only been done in an
handful of cases, mostly focusing on non-rotating or slowly-rotating
BH solutions (see
e.g.~\cite{Ferrari:2000ep,Molina:2010fb,Pani:2009wy,Blazquez-Salcedo:2016enn,Blazquez-Salcedo:2017txk,Brito:2018hjh,Franciolini:2018uyq,Cardoso:2018ptl,Tattersall:2018nve,Tattersall:2019nmh,Blazquez-Salcedo:2019nwd,Silva:2019scu,Glampedakis:2019dqh,Blazquez-Salcedo:2020jee,Blazquez-Salcedo:2020caw,Cano:2020cao})
or relying on the eikonal/geometric optics approximation to obtain
estimates of the QNMs for spinning
BHs~\cite{Blazquez-Salcedo:2016enn,Glampedakis:2017dvb,Jai-akson:2017ldo}. The
only exceptions to this rule, that we are aware of, are computations
of the QNM of Kerr-Newman BHs in Einstein-Maxwell
theory~\cite{Pani:2013ija,Pani:2013wsa,Mark:2014aja,Dias:2015wqa} and
estimates of the BBH ringdown in beyond-GR theories obtained through a
limited number of numerical
simulations~\cite{Okounkova:2019dfo,Okounkova:2019zjf}. Given these
limitations, our ability of going beyond a null test of GR and use our
results to impose precise constraints on beyond-GR theories with QNM
measurements is at this moment quite limited.

Despite these limitations there has been some recent effort to develop
parametrizations that could help mapping measurements of the
parameters $(\delta f_{\ell m 0}, \delta \tau_{\ell m0})$ onto
constraints to specific beyond-GR theories. This includes for example
the parametrization proposed in Ref.~\cite{Maselli:2019mjd}, recently
applied to LIGO-Virgo GW events in Ref. ~\cite{Carullo:2021dui}, where
deviations from the GR QNMs explicitly depend on a perturbative
expansion in the BH spin and possible extra non-GR parameters, or the
proposal of Refs.~\cite{Cardoso:2019mqo,McManus:2019ulj} where
deviations from the GR QNMs are mapped onto generic small
modifications of the perturbation equations describing the QNMs. Other
examples also include proposals to map deviations from the GR QNMs to
coefficients in generic effective field theory
actions~\cite{Cardoso:2018ptl,Franciolini:2018uyq,Cano:2020cao} or to
directly relate measurements of the QNM complex frequencies to a
parametrized non-GR BH
metric~\cite{Glampedakis:2017dvb,Suvorov:2021amy,Volkel:2020daa} which
could be then used jointly with measurements coming from the Event
Horizon Telescope to obtain stronger constraints on deviations from
GR~\cite{Volkel:2020daa,Volkel:2020xlc,Psaltis:2020lvx,Yang:2021zqy}.
We should note, however, that all these parametrizations are either
limited to non-spinning BHs or make use of a series expansion in the
BH spin which might limit their accuracy for highly-spinning BHs.

In the future, it would also be important to test whether this model
could be used to detect deviations in waveforms obtained through
numerical relativity simulations of specific beyond-GR theories. Such
simulations are still at their infancy and have so far only been done
for a handful of theories, focusing mostly on proof-of-concept
simulations~\cite{Healy:2011ef,Berti:2013gfa,Cao:2013osa,Okounkova:2017yby,Hirschmann:2017psw,Witek:2018dmd,Okounkova:2019dfo,Okounkova:2019zjf,Okounkova:2020rqw,East:2020hgw}. Nonetheless,
given the recent efforts put forward in order to simulate BBHs in
beyond-GR theories, we hope that accurate non-GR IMR waveforms will
become available in the near future.

Finally, an obvious generalization of this work would be to extend the
parameterized IMR model to generic precessing BBHs, which can in
principle be easily done using the recently developed multipolar EOB
model reported in Ref.~\cite{Ossokine:2020kjp}.

>>>>

\ab{These constraints could in principle be used to constrain specific
non-GR theories and exotic compact objects~\cite{Glampedakis:2017cgd,Cardoso:2019rvt,Maggio:2020jml}. 
However, QNM computations in non-GR theories have only been done in an
handful of cases, mostly focusing on non-rotating or slowly-rotating
BH solutions~\cite{Ferrari:2000ep,Molina:2010fb,Pani:2009wy,Blazquez-Salcedo:2016enn,Blazquez-Salcedo:2017txk,Brito:2018hjh,Franciolini:2018uyq,Cardoso:2018ptl,Tattersall:2018nve,Tattersall:2019nmh,Blazquez-Salcedo:2019nwd,Silva:2019scu,Glampedakis:2019dqh,Blazquez-Salcedo:2020jee,Blazquez-Salcedo:2020caw,Cano:2020cao}, 
or relying on the eikonal/geometric optics approximation to obtain
estimates of the QNMs for spinning BHs~\cite{Blazquez-Salcedo:2016enn,Glampedakis:2017dvb,Jai-akson:2017ldo}. The
only exceptions to this rule, that we are aware of, are computations
of the QNM of Kerr-Newman BHs in Einstein-Maxwell
theory~\cite{Pani:2013ija,Pani:2013wsa,Mark:2014aja,Dias:2015wqa} or 
estimates of the BBH in non-GR theories obtained through a limited number of 
NR simulations~\cite{Okounkova:2019dfo,Okounkova:2019zjf}. Given these
limitations, our ability of going beyond a null test of GR and use our
results to impose precise constraints on non-GR theories with QNM
measurements is currently quite limited.}

Despite these \ab{theoretical} limitations there has been some recent effort to develop
parametrizations that could help mapping measurements of the
parameters $(\delta f_{\ell m 0}, \delta \tau_{\ell m0})$ onto
constraints to specific non-GR theories. This includes for example
the parametrization proposed in Ref.~\cite{Maselli:2019mjd}, recently
applied to LIGO-Virgo GW events in Ref.~\cite{Carullo:2021dui}, where
deviations from the GR QNMs explicitly depend on a perturbative
expansion in the BH spin and possible extra non-GR parameters, or the
proposal of Refs.~\cite{Cardoso:2019mqo,McManus:2019ulj} where
deviations from the GR QNMs are mapped onto generic small
modifications of the perturbation equations describing the QNMs. Other
examples also include proposals to map deviations from the GR QNMs to
coefficients in generic effective--field-theory
actions~\cite{Cardoso:2018ptl,Franciolini:2018uyq,Cano:2020cao} or to
directly relate measurements of the QNM complex frequencies to a
parametrized non-GR BH metric~\cite{Glampedakis:2017dvb,Suvorov:2021amy,Volkel:2020daa}, 
which could be then used jointly with measurements from the Event
Horizon Telescope to obtain stronger constraints on deviations from
GR~\cite{Volkel:2020daa,Volkel:2020xlc,Psaltis:2020lvx,Yang:2021zqy}.
We should note, however, that all these parametrizations are either
limited to non-spinning BHs or make use of a series expansion in the
BH spin which might limit their accuracy for highly-spinning BHs\ab{, unless 
the sensitivity of GW detectors will not allow us to access the higher 
coefficients in the spin series.}

In the future, it would be important to test whether the $\pSEOB$  model
could be used to detect deviations in waveforms obtained through NR 
simulations of specific non-GR theories. Such results are still at their 
infancy and have so far only been done for a handful of theories, focusing mostly on proof-of-concept
simulations~\cite{Healy:2011ef,Berti:2013gfa,Cao:2013osa,Okounkova:2017yby,Hirschmann:2017psw,Witek:2018dmd,Okounkova:2019dfo,Okounkova:2019zjf,Okounkova:2020rqw,East:2020hgw}. Nonetheless,
given the recent efforts put forward in order to simulate BBHs in
non-GR theories, we hope that accurate non-GR IMR waveforms will become available in the near future.

\ab{Finally, an obvious generalization of this work would be to extend the
parameterized $\pSEOB$ model to generic precessing BBHs, which can in
principle be easily done using the recently developed multipolar EOB
model reported in Ref.~\cite{Ossokine:2020kjp}. Lastly, it will be relevant 
to include GR deviations in the $\pSEOB$ model also during the late inspiral 
and plunge stages of the BBH coalescence. On the other hand, GR deviations (notably 
deviations from PN theory) for the long inspiral stage are currently 
available in the $\pSEOB$ model and have been used to set bounds on the 
PN parameters in the GW phasing using LIGO and Virgo observations~\cite{Abbott:2018lct,
LIGOScientific:2019fpa,Abbott:2020jks}.}
